---
title: "Master Sheet"
author: "Jordan Lo"
date: "2022-11-21"
output: html_document
---
```{r}
# Enables us to read excel files
library("readxl")
# Allows us to do data processing and clean-up
library("tidyverse")
# Enables us to write to an excel file
library("writexl")
```

These chunks of code clean up the data by removing standards and QC as well as any column renaming that needs to happen.

```{r}
# assigns to variables all the data sheets from ang_2, il-18, ngal, and sfas biomarker assays
ang_2 <- read_excel("raw_data\\CHR_US_V1_Ang-2_MSD_Plates1-6_Data_NO_SKS_JH_4-7-2022.xlsx", 
                    sheet = "WideData")
# changes the column name
colnames(ang_2)[2] <- "urine_ang-2_V1 (pg/mL)"

# replaces NaN with NA
ang_2$`urine_ang-2_V1 (pg/mL)`[is.nan(ang_2$`urine_ang-2_V1 (pg/mL)`)] <- NA

# converts the variable type into numeric
ang_2$`urine_ang-2_V1 (pg/mL)` <- as.numeric(as.character(ang_2$`urine_ang-2_V1 (pg/mL)`))
```


```{r}
sfas <- read_excel("raw_data\\CHR_US_V1_sFAS_MSD_Plates1-6_Data_NO_SKS_4-6-2022.xlsx", sheet = "WideSampleData")
sfas <- sfas[-1]
colnames(sfas)[4] <- "urine_sfas_V1 (pg/mL)"
```


```{r}
il18 <- read_excel("raw_data\\CHR_US_V1_IL-18_MSD_Plates1-6_Data_NO__SKS_3-15-2022.xlsx", sheet = "WideData")
il18 <- il18[-1]
colnames(il18)[2] <- "urine_il18_V1 (pg/mL)"

# removes all rows from il18 that are not pure numeric (removing QC/Standard)
il18 <- il18[!is.na(as.numeric(il18$Sample)), ]
```

```{r}
ngal <- read_excel("raw_data\\CHR_US_V1_NGAL_MSD_Plates1-6_Data_NO__SKS_3-16-2022.xlsx", sheet = "WideData")
ngal <- ngal[-1]
colnames(ngal)[2] <- "urine_ngal_V1 (pg/mL)"

ngal <- ngal[!is.na(as.numeric(ngal$Sample)), ]
```

This code chunk merges the assay data for all four of the biomarkers and then orders them by sample id
```{r}
# does a left join of the sfas and ang_2 dataframes by "Sample"
master <- merge(sfas, ang_2, by = "Sample")

# does a partial join of the datasets
master <- merge(master, il18, by = "Sample", all.y = TRUE)
master <- merge(master, ngal, by = "Sample", all.y = TRUE)

# orders the dataframe by Sample number
master <- master[order(master$Sample),]
```

```{r}
# creates a new excel sheet of the master dataframe with the assay results
write_xlsx(master, "data_merges\\assay_master_sheet.xlsx")
```

Here, I'm making Pavan's requested table with all of the variables he wanted, and then exporting it as an xlsx
```{r}
# creating a new dataframe
summary <- data.frame(urine_biomarker = c('NGAL', 'IL-18', 'sFas', 'Ang_2'), 
                      N = c(202, 202, 202, 202), 
                      number_LLOD = c(0, 0, 0, 149), 
                      avg_interplate_cv = c(4.7, 2.9, 2.5, 18.7), 
                      avg_intraplate_CV = c(3.4, 2.8, 2.5, 19.2), 
                      avg_concentration_pg_per_ml = c(1390947, 1903.73, 7277, 1059.806), 
                      iqr = c(569731-58871, 945.45-214.65, 8720-2767, 11013.661-43.253)
                      )
write_xlsx(summary, "data_merges\\biomarker_summary_20221122.xlsx")
```

Now I'm going to be merging the master dataframe from above with some clinical data that I was given
```{r}
# pulling a csv and storing it under a variable
incident <- read_csv("raw_data\\chrome_incident_aki_data_for_neha_10_26_2022.csv")

prevalent <- read_csv("raw_data\\chrome_prevalent_aki_for_neha_10_26_2022.csv")

aki_data <- merge(incident, prevalent, by = "Subject_ID")

colnames(aki_data)[1] <- "StudyID"
```

```{r}
biomarker_assays_clinical_data <- merge(master, aki_data, by = "StudyID", all.x= TRUE)

write_xlsx(biomarker_assays_clinical_data, "data_merges\\Biomarker_Assays_and_Clinical_Data_20221122.xlsx")
```
```{r}
bio_clinical <- read_excel("data_merges\\Biomarker_Assays_and_Clinical_Data_20221122.xlsx")

# creates new columns for the log transformations of the biomarker concentrations 
bio_clinical$log_ang2 <- log(bio_clinical$`urine_ang-2_V1 (pg/mL)`)
bio_clinical$log_il18 <- log(bio_clinical$`urine_il18_V1 (pg/mL)`)
bio_clinical$log_ngal <- log(bio_clinical$`urine_ngal_V1 (pg/mL)`)
bio_clinical$log_sfas <- log(bio_clinical$`urine_sfas_V1 (pg/mL)`)

# creates a pairs plot of the biomarker concentration data using ggpairs
bio_clinical %>%
  select(c(`log_ang2`, `log_il18`, `log_ngal`, `log_sfas`, )) %>%  # selects the data used
  ggpairs(labels = letters[1:ncol(bio_clinical)],                  # creates the actual plot
          lower = list(continuous=wrap("smooth"))) +               
  theme_bw() +                                                     # color scheme of plot
  ggtitle("Correlation Data of Biomarker Concentrations") +        # creates a title
  theme(plot.title = element_text(hjust = 0.5))                    # centers title

ggsave(file = "visualizations\\biomarker_pairs.png",
       height = 10,
       width = 14)
```


